---
title: "HW 4"
output:    
  github_document:
    html_preview: false
  html_document: default
  word_document: default
always_allow_html: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(microbenchmark)
library(timeSeries)
library(parallel)
```
### HPC:
#### Problem 1:
```{r Total row sums}
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

# Rewrite
fun1alt <- function(mat) {
  ans<-rowSums(mat)
  ans
}
```

```{r Cumulative sum by row}
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
   ans <- rowCumsums(mat)
   ans
}
```

```{r}
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), unit = "relative", check = "equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), unit = "relative", check = "equivalent"
)
```

#### Problem 2: Make things run faster with parallel computing
```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132

# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})


# Rewrite the previous code using parLapply() to make it run faster. 
cl <- makePSOCKcluster(4L) 
clusterSetRNGStream(cl, 123)
clusterExport(cl, c("sim_pi"), envir = environment())

system.time({
  ans <- unlist(parLapply(cl,1:4000, sim_pi, n = 10000)) 
  print(mean(ans))
})

stopCluster(cl)
```
### SQL:
```{r}
library(RSQLite)
library(DBI)
```

```{r Setup}
# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```

#### Question 1: How many movies is there avaliable in each rating catagory.
```{sql connection=con}
SELECT rating,
  COUNT(*) AS count
FROM film
GROUP BY rating
```
There are 180 movies with rating "G"; 210 movies with rating "NC-17"; 194 movies with rating "PG"; 233 movies with rating "PG-13"; 195 movies with rating "R".

#### Question 2: What is the average replacement cost and rental rate for each rating category.
```{sql connection=con}
SELECT rating,
  AVG(replacement_cost) AS avg_replacement_cost, 
  AVG(rental_rate) AS avg_rental_rate
FROM film
GROUP BY rating
```
For the category with rating "G", the average replacement cost is 20.12 and average rental rate is 2.91;For the category with rating "NC-17", the average replacement cost is 20.13 and average rental rate is 2.97;For the category with rating "PG", the average replacement cost is 18.96 and average rental rate is 3.05;For the category with rating "PG-13", the average replacement cost is 20.40 and average rental rate is 3.03;For the category with rating "R", the average replacement cost is 20.23 and average rental rate is 2.94.

#### Question 3: Use table film_category together with film to find the how many films there are witth each category ID.
```{sql connection=con}
Select category_id,
  COUNT(*) AS count
FROM film AS a INNER JOIN film_category AS b 
  ON a.film_id = b.film_id
GROUP BY category_id
```
#### Question 4: Incorporate table category into the answer to the previous question to find the name of the most popular category.
```{sql connection=con}
SELECT film_category.category_id AS Category_id, category.name AS Name,
  COUNT(*) AS count
FROM film_category
  INNER JOIN film ON film_category.film_id = film.film_id
  INNER JOIN category ON category.category_id = film_category.category_id
GROUP BY film_category.category_id
ORDER BY COUNT (*) DESC 
LIMIT 3
```
The most popular category is sports which contains 74 movies. 














