---
title: "LAB 7"
output:    
  github_document:
    html_preview: false
  html_document: default
  word_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
options(repos = c(CRAN = "http://cran.rstudio.com"))
```

### API
```{r}
library(tidyverse)
library(tidytext)
library(httr)
library(xml2)
library(stringr)
library(tibble)
library(ggplot2)
```
#### Step 1:Using the NCBI API, look for papers that show up under the term “sars-cov-2 trial vaccine” Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]/span")

# Turning it into text
counts <- as.character(counts)

# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
```
I founde 2336 papers under the term "sars-cov-2 trial vaccine".

#### Step 2:Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.
```{r}
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
              db = "pubmed",
              term = "sars-cov-2 trial vaccine",
              retmax = 250)
)

# Extracting the content of the response of GET
ids <- httr::content(query_ids)
ids

# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "<Id>|</Id>")


publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db = "pubmed",
    id =I(paste(ids,collapse=",")),
    retmax = 1000,
    rettype = "abstract"
    )
)

# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```

#### Step 3:As we did in lab 7. Create a dataset containing the following: Pubmed ID number, Title of the paper, Name of the journal where it was published, Publication date, and Abstract of the paper (if any).
```{r}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```
#### Geitting the abstracts, title, name of journal and date
```{r}
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>") 
abstracts <- str_replace_all(abstracts, "[[:space:]]+", " ")

titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")
 
date <- str_extract(pub_char_list, "<PubDate>[[:print:][:space:]]+</PubDate>")
date <- str_remove_all(date, "</?[[:alnum:]- =\"]+>")
date <- str_replace_all(date, "[[:space:]]+", " ")

name <- str_extract(pub_char_list, "<Title>[[:print:][:space:]]+</Title>")
name <- str_remove_all(name, "</?[[:alnum:]- =\"]+>")
name <- str_replace_all(name, "[[:space:]]+", " ")

# Make a table
database <- data.frame(
   PubMedId = ids,
   Title    = titles,
   Journal  = name,
   date     =date,
   Abstract = abstracts
 )
knitr::kable(database[1:20,], caption = "Papers about sars-cov-2 trial vaccine")
```

### Text Mining

#### Step 1:Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?
```{r}
fn<-"pubmed.csv"
if (!file.exists(fn))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv", destfile = fn)

pubmed<-read.csv(fn)
pubmed<-as.tibble(pubmed)

# Token the abstracts and count the number of each token
pubmed %>%
  unnest_tokens(output=word, input=abstract) %>%
  count(word, sort = TRUE) %>%
  top_n(20) %>%
  ggplot(aes(x=n, y=fct_reorder(word, n))) +
  geom_col()

# Remove stop words
pubmed %>%
  unnest_tokens(output=word, input=abstract) %>%
  count(word, sort = TRUE) %>%  
  anti_join(stop_words, by="word")%>%
  # Using regular expression to remove numbers
  filter(!grepl("^[0-9]+$", x= word))%>%
  top_n(20) %>%
  ggplot(aes(x=n, y=fct_reorder(word, n))) +
  geom_col()

# 5 most tokens for each search
pubmed %>% 
  unnest_tokens(output=word, input=abstract)%>%
  group_by(term)%>%
  count(word)%>%
  anti_join(stop_words, by="word")%>%
  top_n(5,n)%>%
  knitr::kable()
```
Before removing the stop words, the most common 5 words are "the", "of", "and", "in" and "to" which are all stop words and meaningless to collect tokens like this. After removing stop words and numbers, we got the most common 5 words are "covid", "patients", "cancer", "prostate" and "disease" which are different from before removing stop words.The most common 5 tokens for "covid": "19","covid","disease","pandemic","patients"; most common 5 tokens for cystic fibrosis are: "cf", "cystic", "disease", "fibrosis", "patients"; the most common 5 tokens for meningitis are: "clinical", "csf", "meningeal", "meningitis", "patients";  the most common 5 tokens for meningitis preeclampsia are: "eclampsia", "pre", "preeclampsia", "pegnancy”,“women”；the most common 5 tokens for meningitisprostate cancer are: "cancer", "disease", "patients", "prostate", "treatment".

#### Step 2:Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.
```{r}
# Bigram
pubmed %>%
  unnest_ngrams(output=bigram, input=abstract, n=2) %>%
  count(bigram, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x=n, y=fct_reorder(bigram, n))) +
  geom_col()
```
The most 10 biagrams are ’covid 19‘, ‘of the’, ‘in the’, ‘prostate cancer’, ‘pre eclampsie’, ‘patients with’, ‘of covid’, ‘and the’, ‘to the’, and ‘of prostate’.

#### Step 3:Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the “document”) What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?
```{r}
pubmed %>%
  unnest_tokens(word, abstract) %>%
  count(word, term) %>%
  bind_tf_idf(word, term, n) %>%
  group_by(term)%>%
  top_n(5,tf_idf)%>%
  arrange(desc(tf_idf), .by_group = TRUE)%>%
  knitr::kable()
```
The most common 5 tokens for "covid": "covid","sars","pandemic","coronavirus", "cov"; most common 5 tokens for cystic fibrosis are: "cf", "cystic", "sweat", "fibrosis", "cftr"; the most common 5 tokens for meningitis are: "meninges", "csf", "meningeal", "meningitis", "pachymeningitis	";  the most common 5 tokens for meningitis preeclampsia are: "eclampsia", "maternal", "preeclampsia", "pegnancy”,“gestational”；the most common 5 tokens for meningitisprostate cancer are: "castration", "prostatectomy", "psa", "prostate", "androgen".

Compared to Question 1, For the term "COVID" there are 2 different word-search term combination occurred. For the term "cystic fibrosis" there are 2 different word-search term combination occurred. For the term "meningitis" there are 2 different word-search term combination occurred. For the term "preeclampsia" there are 2 different word-search term combination occurred.For the term "	prostate cancer" there are 4 different word-search term combination occurred. 
